The SMT Competition 2015--2018

### 2019-08-27 2018/report/Reviews_67398_Revised_Response.txt

> *** Review 1 ***
>
>  The paper overall improved since the first submission, but there are still issues to be fixed, some of them already reported in the review of the first version.
>
>  * Organization
>
> - information about ranking should all stay in the related section, while
> there are related information, e.g. in the first part of Section 7 - Results.

DONE (all information about the scoring rules was moved into a new
Section 7)

@@

> Fig 3/4: Starting the X-axis at an arbitrary point is always a bit fishy. This
> way, any tiny difference can be overemphasized to achieve any level of
> impressiveness that the author choses. For instance, 144k vs 136k instances
> (casually read off the plot) is a 3.5% increase, but this is an irrelevant
> detail as explained in the text. All four series are also of exactly the same
> shape. I think a cactus plot is simply not the right tool for this job.

WONTFIX (starting the x-axis at 0 would render the plots unexpressive;
a cactus plot may not be ideal but we haven't been able to find a
better tool)

> ============================================
>
> *** Review 3 ***
>
>
> This is a revised version of a paper describing
> the years 2015-2018 of the SMT competition.
> The important changes requested last time were addressed,
> and new content was added in Section 8,
> which discusses interesting questions
> and adds substantial value to the paper.
>
> There are only a few minor improvements that should be made before publication:
>
> In the definition of the raw benchmark score (Sect. 4.1),
> it is not clear why CPU time is bounded at 4T
> (the fact that only 4 cores are available only comes later).

DONE (raw benchmark scores are now introduced in a new Section 7)

@@

> In the remainder of Sect. 4.1, the term "benchmark family" is used,
> but this is never defined.
> This could be added to the definitions part in the beginning of Sect. 4.

DONE (benchmark families are now introduced in a new Section 7)

@@

> The first paragraph of Sect. 8 mentions only three aspects,
> but more aspects are analyzed.

DONE

> Figure 5 seems unnecessary, because the difference is not visible anyway.
> Just listing the numbers would be sufficient.

WONTFIX (we have a similar figure for each track; omitting it for the
Application Track merely for space reasons seems unnecessary, and it
serves to show that the difference is hardly visible---however, we
reduced the size of the figure)

### 2019-08-27 2018/report/SMT-COMP.tex
Scoring

\paragraph{Eligible Benchmarks.}
Some benchmarks typically have to be
excluded for syntactic reasons: in 2015 and 2016, \num{11966} benchmarks
were excluded because they contained partial or underspecified
operations (e.g., bit-vector division), whose precise semantics were
being debated at the time.  Additionally in~2016, one benchmark was
excluded because it contained incorrect status information.  In~2017,
\num{1104} benchmarks were excluded because their logic had been classified
incorrectly, and 8~benchmarks were excluded because of incorrect or
missing status information.  Moreover, divisions containing datatypes
({DT})---which were new in~2017---were excluded from the Unsat-Core
track because of inadequate tool support.  In~2018, \num{3603}~benchmarks were
excluded because they contained named terms (a feature of the SMT-LIB
language that is not permitted in competition benchmarks), and
5~benchmarks were excluded because of missing status information.

@@

\paragraph{Benchmark Selection.}
SMT Competitions until~2014 employed a pseudo-random selection process
to further cull the number of benchmarks used in the competition.
However, a 2013 evaluation of SMT-COMP and SMT-LIB found that this
``significantly lessens the ability of a competition to determine
`best' solvers''~\cite{CSW15}. Together with the increased
computational resources available to the competition since its move to
StarExec, this finding prompted a paradigm shift.  Since 2015, the
competition is evaluating all solvers on all eligible benchmarks.

\paragraph{Benchmarks with Unknown Status.}
Another paradigm shift concerns the treatment of benchmarks with
unknown status.  Because of the difficulty to judge solver responses
for such benchmarks, SMT Competitions until 2015 only used benchmarks
whose status was known (i.e., that had been solved previously by
existing SMT solvers).  In 2015, \num{30717} non-incremental benchmarks
from the SMT Library were not used for the competition because their
status was unknown.  However, this approach was criticized for
imitating existing solvers rather than truly innovating.  In~2016,
the SMT Library contained \num{29724} non-incremental benchmarks with
unknown status; solver performance on these was evaluated but reported
separately.  Since 2017, all non-incremental benchmarks are eligible
for the competition's \maintrack regardless of status, and they
directly affect the competition results.  Solver responses for such
benchmarks are assumed to be correct, unless there is a disagreement
between two or more solvers that are otherwise (i.e., on benchmarks with
known status) sound.  Benchmarks with disagreements are removed from
the competition results.  This conservative approach was chosen over a
voting system as the latter is unreliable (especially in divisions
with few solvers), could lead to a solver being punished wrongly, and
could be gamed by entering multiple copies of a solver into the
competition.  Moreover, the size and time frame of the competition
demand an automated resolution.  Disagreements are reported to the
solver developers involved, who typically perform a manual
investigation at their leisure, and in all cases have led to bugs in
solvers being identified and fixed.  In 2016, there were disagreements
on 79 benchmarks; in 2017, only one benchmark was removed; while in
2018, there were no disagreements between otherwise sound solvers.

@@

@@

Solvers in each division (of each track) are ranked according to a
metric that is based, first, on the number of erroneous responses
(which is usually~0, but solvers that did give erroneous responses are
ranked below solvers that did not); second, on the number of correct
responses; third and fourth, on the wall-clock and CPU time,
respectively, that the solver process consumed.  For the \ucoretrack
(which uses unsatisfiable benchmarks only), the reduction in the size
of the unsatisfiable core is used instead of the number of correct
responses.  A detailed description of the scoring can be found in the
competition rules~\cite{rules18}.

The \emph{raw benchmark score} of a solver is thus a quadruple
$\langle e, n, w, c\rangle$, with
\begin{itemize}[noitemsep]
  \item \makebox[5em][l]{$e\geq 0$} number of erroneous responses (usually~$e = 0$),
  \item \makebox[5em][l]{$0 \leq n \leq N$} number of correct responses (resp.~\emph{reduction} for the \ucoretrack),
  \item \makebox[5em][l]{$w \in [0,T]$} wall-clock time in seconds (real-valued),
  \item \makebox[5em][l]{$c \in [0, 4T]$} CPU time in seconds (real-valued).
\end{itemize}
%
For the \maintrack and \apptrack, $e$ is the number of returned
statuses that disagree with the given expected status; $e \in \{0,1\}$
for the \maintrack.  (As described in Section~\ref{sec:benchmarks},
disagreements between otherwise sound solvers on benchmarks with
unknown status lead to the benchmark being disregarded.)  For the
\ucoretrack, $e$ includes, in addition, the number of returned unsat
cores that are ill-formed or are not, in fact, unsatisfiable (as
validated by a selection of other solvers selected by organizers).

For the \maintrack and \apptrack,
$N$ is defined as the number of {\tt check-sat} commands ($N=1$ for the \maintrack), and
$n$ is defined as the number of correct results.
For the \ucoretrack, $N$ is defined as the number of named top-level assertions,
and~$n$ is defined as the \emph{reduction}, i.e., the difference between $N$
and the size of the unsat core returned by the solver.

This raw score corresponds to the \emph{raw parallel score} (all values are
determined for solver performance within the wall-clock time limit).  To
evaluate sequential performance, we derive a \emph{raw sequential score} by
imposing a (virtual) CPU time limit equal to the wall-clock time limit~$T$.  A
solver result is taken into consideration for the sequential score only if the
solver process terminates within this CPU time limit.

@@

All solvers are run with four cores available, and the time limit~$T$
that is imposed on each job pair (see Table~\ref{table:time-limits})
is a wall-clock limit.  Thus, solvers that take advantage of
parallelism might use up to~$4T$ CPU time.  For the \maintrack and the
\ucoretrack, the competition reports parallel and sequential
performance separately.  The latter is determined by imposing a
(virtual) CPU time limit of~$T$.  For the \apptrack, only
parallel performance is reported (because response times to individual
{\tt check-sat} commands are not recorded by StarExec).

@@

Until 2015, the competition used these raw scores for all tracks to
determine the division score of a solver as the sum of the scores of
all benchmarks in that division.  Fewer errors in a division takes
precedence over more correct solutions (resp.~reduction), which takes
precedence over less wall-clock time taken, which takes precedence
over less CPU time taken.  However, since there are vast differences
in size between benchmark families (i.e., groups of benchmarks with
similar characteristics) in the SMT Library, the competition
since~2016 uses a scoring method that assigns a weight to each
benchmark in the \maintrack and \ucoretrack in order to de-emphasize
large benchmark families~\cite{rules16,rules17,rules18}.  For
application track divisions, division scores were still based on raw
scores only, since application track benchmarks may be partially
solved.  Defining a useful sequential score for the application track
would require information not provided by the raw score, e.g.,
detailed timing information for each result.

\paragraph{Benchmark Weights (\maintrack and \ucoretrack 2016--2018).} The \emph{weight} of a
benchmark in the main track is computed as $\alpha = (1 + \ln F)/F$,
where~$F\geq 1$ is the size of the benchmark's family.  We then define its
\emph{normalized weight} as $\alpha' = \alpha / (\sum_{i\in F} \alpha_i)$,
where the sum is over all benchmarks in the division.  The log scaling seems a
reasonable compromise between linearly combining numbers of benchmarks, which
would overemphasize large families, and simply averaging the fraction of
benchmarks solved, which would overemphasize small families.

\paragraph{Weighted Division Scores (\maintrack and \ucoretrack 2016--2018).}
The \emph{(weighted) score} of a solver in a division of the \maintrack and
\ucoretrack with $M$ benchmarks is then defined as the weighted sum of all raw
scores $$\sum_{b\in M} \alpha'_b \cdot \langle e_b \cdot N, n_b \cdot N,
w_b, c_b\rangle$$ where the sum is over all benchmarks in the division to
assess parallel performance, and the weighted sum of all sequential scores to
assess sequential performance.

\subsection{Competition-Wide Ranking}
\label{sec:scoring-floc}

@@

The competition-wide ranking for the \maintrack was retained and
refined after~2014.  Table~\ref{table:competition-wide-ranking} shows
the best three solvers according to this ranking for 2015--2018.  Also
shown is Z3, which ranked in high position each year but was running
hors concours.  Sequential and parallel performance are reported
separately.

\paragraph{Competition-Wide Score (\maintrack).}
A competition-wide metric for the main track, separately for
parallel and sequential performance, was defined as follows.

Let $N_i$ be the total number of benchmarks in division~$i$ that were used in the
competition (and not removed because of disagreements), and let $\langle e_i, n_i, w_i, c_i\rangle$ be a solver's
raw score for this division.
%
In 2015 and 2016,
the \emph{competition-wide raw score} for a solver was defined as
%
$\sum_i (e_i == 0 \,?\, (n_i/N_i)^2 : -e_i) \log N_i$,
%
where the sum is over all competitive
divisions into which the solver was
entered.
%
In 2017, the (per-division) penalty for erroneous results was changed from the
number of errors to a fixed constant, since the weight already takes the number
of benchmarks in the division into account, and multiplying it by the number of
errors overemphasizes large divisions.
In 2017 and 2018,
the \emph{competition-wide raw score} for a solver was thus defined as
%
$\sum_i (e_i == 0 \,?\, (n_i/N_i)^2 : -4) \log_e N_i$,
%
where the sum is over all competitive divisions into which the solver
was entered.

A contentious issue with this ranking has been the scoring of
incorrect solver responses.  In the division rankings, an incorrect
response causes a solver to be ranked below all correct solvers for
that division.  But the competition's \maintrack consists of many
independent divisions.  A solver bug that affects one division need
not manifest itself in other divisions.  To balance the strong
interest in correct solvers against the risk of stifling innovation,
the competition-wide ranking was revised in~2017 to use a fixed
numeric penalty for each division with incorrect responses.  The value
of this penalty is currently set so that entering a (possibly buggy)
solver that can solve all benchmarks into a division has a positive
expected score if the probability that a soundness bug in the solver
will be triggered is below~$20\,\%$.~\cite{rules18}

>>>>

\paragraph{Eligible benchmarks.}
Some benchmarks typically have to be
excluded for syntactic reasons: in 2015 and 2016, \num{11966} benchmarks
were excluded because they contained partial or underspecified
operations (e.g., bit-vector division), whose precise semantics were
being debated at the time.  Additionally in~2016, one benchmark was
excluded because it contained incorrect status information.  In~2017,
\num{1104} benchmarks were excluded because their logic had been classified
incorrectly, and 8~benchmarks were excluded because of incorrect or
missing status information.  Moreover, divisions containing datatypes
({DT})---which were new in~2017---were excluded from the Unsat-Core
track because of inadequate tool support.  In~2018, \num{3603}~benchmarks were
excluded because they contained named terms (a feature of the SMT-LIB
language that is not permitted in competition benchmarks), and
5~benchmarks were excluded because of missing status information.

@@

\paragraph{Benchmark selection.}
SMT Competitions until~2014 employed a pseudo-random selection process
to further cull the number of benchmarks used in the competition.
However, a 2013 evaluation of SMT-COMP and SMT-LIB found that this
``significantly lessens the ability of a competition to determine
`best' solvers''~\cite{CSW15}. Together with the increased
computational resources available to the competition since its move to
StarExec, this finding prompted a paradigm shift.  Since 2015, the
competition is evaluating all solvers on all eligible benchmarks.

\paragraph{Benchmarks with unknown status.}
Another paradigm shift concerns the treatment of benchmarks with
unknown status.  Because of the difficulty to judge solver responses
for such benchmarks, SMT Competitions until 2015 only used benchmarks
whose status was known (i.e., that had been solved previously by
existing SMT solvers).  In 2015, \num{30717} non-incremental benchmarks
from the SMT Library were not used for the competition because their
status was unknown.  However, this approach was criticized for
imitating existing solvers rather than truly innovating.  In~2016,
the SMT Library contained \num{29724} non-incremental benchmarks with
unknown status; solver performance on these was evaluated but reported
separately.  Since 2017, all non-incremental benchmarks are eligible
for the competition's \maintrack regardless of status, and they
directly affect the competition results.  Solver responses for such
benchmarks are assumed to be correct, unless there is a disagreement
between two or more solvers that are otherwise (i.e., on benchmarks with
known status) sound.  Benchmarks with disagreements are removed from
the competition results.  This conservative approach was chosen over a
voting system as the latter is unreliable (especially in divisions
with few solvers), could lead to a solver being punished wrongly, and
could be gamed by entering multiple copies of a solver into the
competition.  Moreover, the size and time frame of the competition
demand an automated resolution.  Disagreements are reported to the
solver developers involved, who typically perform a manual
investigation at their leisure, and in all cases have led to bugs in
solvers being identified and fixed.  In 2016, there were disagreements
on 79 benchmarks; in 2017, only one benchmark was removed; while in
2018, there were no disagreements between otherwise sound solvers.

@@

A detailed description of the scoring can be found in the competition
rules~\cite{rules18}.  We summarize the most important aspects in this
section.

@@

Solvers in each division (of each track) are ranked according to a
metric that is based, first, on the number of erroneous responses
(which is usually~0, but solvers that did give erroneous responses are
ranked below solvers that did not); second, on the number of correct
responses; third and fourth, on the wall-clock and CPU time,
respectively, that the solver process consumed.
%
The \emph{raw benchmark score} of a solver is thus a quadruple
$\langle e, n, w, c\rangle$, with
\begin{itemize}[noitemsep]
  \item \makebox[5em][l]{$e\geq 0$} number of erroneous responses (usually~$e = 0$),
  \item \makebox[5em][l]{$0 \leq n \leq N$} number of correct responses (resp.~\emph{reduction} for the \ucoretrack),
  \item \makebox[5em][l]{$w \in [0,T]$} wall-clock time in seconds (real-valued),
  \item \makebox[5em][l]{$c \in [0, 4T]$} CPU time in seconds (real-valued).
\end{itemize}
%
For the \maintrack and \apptrack, $e$ is the number of returned
statuses that disagree with the given expected status; $e \in \{0,1\}$
for the \maintrack.  For the \ucoretrack, $e$ includes, in addition,
the number of returned unsat cores that are ill-formed or are not, in
fact, unsatisfiable (as validated by a selection of solvers selected
by the organizers).

For the \maintrack and \apptrack, $N$ is defined as the number of {\tt
  check-sat} commands ($N=1$ for the \maintrack), and $n$ is defined
as the number of correct responses.  For the \ucoretrack (which uses
unsatisfiable benchmarks only), $N$ is defined as the number of named
top-level assertions, and~$n$ is defined as the \emph{reduction},
i.e., the difference between $N$ and the size of the unsat core
returned by the solver.

@@

@@

\paragraph{Parallel vs.\ sequential performance.}

All solvers are run with four cores available, and the time limit~$T$
that is imposed on each job pair (see Table~\ref{table:time-limits})
is a wall-clock limit.  Thus, solvers that take advantage of
parallelism might use up to~$4T$ CPU time.  For the \maintrack and the
\ucoretrack, the competition reports parallel and sequential
performance separately.  The raw benchmark score corresponds to the
\emph{raw parallel score}---all values are determined for solver
performance within the wall-clock time limit.  We derive a \emph{raw
  sequential score} by imposing a (virtual) CPU time limit equal to
the wall-clock time limit~$T$.  A solver response is taken into
consideration for the sequential score only if the solver process
terminates within this CPU time limit.  For the \apptrack, only
parallel performance is reported (because response times to individual
{\tt check-sat} commands are not recorded by StarExec).

\paragraph{Benchmark weights.}

Until 2015, the competition used these raw scores for all tracks to
determine the division score of a solver as the sum of the scores of
all benchmarks in that division.  However, there are vast differences
in size between benchmark \emph{families} (i.e., groups of benchmarks
with similar characteristics) in the SMT Library.  The pseudo-random
selection process of benchmarks that was used until~2014 (see
Section~\ref{sec:benchmarks}) ensured that large families were not
overemphasized in the competition.  Since~2016, the competition uses a
scoring method that assigns a weight to each benchmark in the
\maintrack and \ucoretrack to achieve the same
goal~\cite{rules16,rules17,rules18}.

The \emph{weight} of a benchmark~$b$ in the \maintrack is computed as
$\alpha_b = (1 + \ln F_b)/F_b$, where $F_b \geq 1$ is the size of the
benchmark's family.  We then define its \emph{normalized weight} as
$\alpha'_b = \alpha_b / (\sum_{b'} \alpha_{b'})$, where the sum is
over all benchmarks in the division.  The logarithmic scaling seems a
reasonable compromise between linearly combining numbers of
benchmarks, which would overemphasize large families, and simply
averaging the fraction of benchmarks solved, which would overemphasize
small families.

The \emph{(weighted) score} of a solver in a (\maintrack or
\ucoretrack) division~$D$ is then defined as the weighted sum of all
raw scores
%
$$\sum_{b\in D} \alpha'_b \cdot \langle e_b \cdot |D|, n_b \cdot |D|, w_b,
  c_b\rangle$$
%
where the sum is over all benchmarks in the division, and~$|D|$ is the
number of benchmarks in~$D$.  This is computed separately for parallel
and sequential performance.

For \apptrack divisions, division scores were still based on raw
scores only, since \apptrack benchmarks may be partially solved.

\subsection{Competition-Wide Ranking (\maintrack)}
\label{sec:scoring-floc}

@@

The competition-wide ranking for the \maintrack was retained and
refined after~2014.  Sequential and parallel performance are reported
separately.  Let $N_i$ be the total number of benchmarks in
division~$i$ that were used in the competition (and not removed
because of disagreements), and let $\langle e_i, n_i, w_i, c_i\rangle$
be a solver's raw score for this division.
%
In 2015 and~2016, the \emph{competition-wide raw score} for a solver
was defined as
%
$$\sum_i (e_i == 0 \,?\, (n_i/N_i)^2 : -e_i) \log N_i$$
%
where the sum is over all competitive divisions into which the solver
was entered.
%
In 2017, the (per-division) penalty for erroneous results was changed
from the number of errors to a fixed constant, since the division
weight~$\log N_i$ already takes the number of benchmarks in the
division into account, and multiplying it by the number of errors
overemphasizes large divisions.  In 2017 and 2018, the
\emph{competition-wide raw score} for a solver was thus defined as
%
$$\sum_i (e_i == 0 \,?\, (n_i/N_i)^2 : -4) \log_e N_i$$
%
The value of this constant was chosen to balance the strong interest
in correct solvers against the risk of stifling innovation.  In the
division rankings (Section~\ref{sec:scoring-division-rankings}), an
incorrect response causes a solver to be ranked below all correct
solvers for that division.  But the competition's \maintrack consists
of many independent divisions; a solver bug that affects one division
need not manifest itself in other divisions.  With the current value,
entering a (possibly buggy) solver that can solve all benchmarks into
a division has a positive expected score if the probability that a
soundness bug in the solver will be triggered is below~$20\,\%$.
